a and generators

### 2022-05-03 documentos/proofdiagonalization.tex
comentarios

\subsection{Single qubit} % {{{
\ddnote{Alejo, por fa revise q no se me fue un typo con el asunto de mod 2, ya corregí toda la notación y removi los otros comentarios, que ya fueron aclarados en la reunion y trataban de lo mismo} \afnote{Listo, todo en orden.}

@@

the sake of clarity, we will use \afnote{we use} Latin symbols for binary indices, and

@@

% }}}
\subsection{$N$ qubits} % {{{
To work out the $N$ qubit case, we again rely on binary indices. In this case, we replace 
$N$ dimensional vector $\valpha$ with a pair of $N$ dimensional vector binary
indices $\vec j$ and $\vec k$ so that 
each entry $\alpha_i$ of $\valpha$ is identified with the pair $j_i$ and $k_i$ as in the 
single qubit case of the previous subsection. Then, all the steps leading to 
\eref{eq:lambda:is:tau:appendix} can be redone, but with vector indices instead of binary scalars.

\ddnote{Poner atencion a lo siguiente}
{\red[viejo y mal]The tensor product of Pauli matrices, in vector form,  will be denoted by 
$\kket{\sigma_{\vec{k}\vec{l} }} = \kket{\sigma_{k_1l_1}}\otimes \cdots \otimes \kket{\sigma_{k_Nl_N}}$
and fulfill the orthogonality relation 
$\bbrakket{\sigma_{\vec{k}\vec{l} }}{\sigma_{\vec{k}'\vec{l}'}} = 
2^N \delta_{\vec{k}\vec{k}'}\delta_{\vec{l}\vec{l}'}$.}
{\blue [nuevo y correcto y breve, la condicion de ortogonalidad no se usa explicitamente y es trivial]
The tensor product of Pauli matrices, in vector form,  will be denoted by 
$\kket{\sigma_{\vec{k}\vec{l} }}$.
} \afnote{perfecto}
% \begin{equation}
% \bbrakket{\sigma_{\vec{k}\vec{l} }}{\sigma_{\vec{k}'\vec{l}'}} = 2^N \delta_{\vec{k}\vec{k}'}\delta_{\vec{l}\vec{l}'}.
% \end{equation}
With this in mind, a $N$-qubit Pauli map can be written as 
% 
% 
% Analogously to the previous case, we can define an arbitrary Pauli map acting on $N$ qubits density matrices in terms of the tensor products of vectorized Pauli matrices
\begin{equation}
\hat{\mathcal{E}}_N = 
  \frac{1}{2^N}\sum_{\vec{l}\vec{m}} \tau_{\vec{l}\vec{m}} \ddyada{\sigma_{\vec{l}\vec{m} }}. 
\end{equation}
The generalization of Eqs.~(\ref{CompToPau}) and (\ref{PauToComp}) read 
\ddnote{corregir el angulo doble y tambien el oplus}
\begin{align}
\kket{\sigma_{\vec{k}\vec{l} }}  & = 
   \sum_{\vec{j}} (-1)^{\vec{j}\cdot\vec{k}} \ket{\vec{j}}\ket{\vec{j}+ \vec{l}(\text{mod}2)},\\
%
 \ket{\vec{l}}\ket{\vec{l}+ \vec{n}(\text{mod}2)} & = 
   \frac{1}{2^N}\sum_{\vec{m}} (-1)^{\vec{m}\cdot\vec{l}}\kket{\sigma_{\vec{m}\vec{n} }}.
\end{align}
% \afnote{Cambiar los $\mu$ y $\nu$ por indices latinos.}

>>>>

@@

the sake of clarity, we use Latin symbols for binary indices, and

@@

% }}}
\subsection{$N$ qubits} % {{{
To work out the $N$ qubit case, we again rely on binary indices. In this case, we replace 
$N$ dimensional vector $\valpha$ with a pair of $N$ dimensional vector binary
indices $\vec j$ and $\vec k$ so that 
each entry $\alpha_i$ of $\valpha$ is identified with the pair $j_i$ and $k_i$ as in the 
single qubit case of the previous subsection. Then, all the steps leading to 
\eref{eq:lambda:is:tau:appendix} can be redone.

The tensor product of Pauli matrices, in vector form,  will be denoted by 
$\kket{\sigma_{\vec{k}\vec{l} }}$.
% \begin{equation}
% \bbrakket{\sigma_{\vec{k}\vec{l} }}{\sigma_{\vec{k}'\vec{l}'}} = 2^N \delta_{\vec{k}\vec{k}'}\delta_{\vec{l}\vec{l}'}.
% \end{equation}
With this in mind, a $N$-qubit Pauli map can be written as 
% 
% 
% Analogously to the previous case, we can define an arbitrary Pauli map acting on $N$ qubits density matrices in terms of the tensor products of vectorized Pauli matrices
\begin{equation}
\hat{\mathcal{E}}_N = 
  \frac{1}{2^N}\sum_{\vec{l}\vec{m}} \tau_{\vec{l}\vec{m}} \ddyada{\sigma_{\vec{l}\vec{m} }}. 
\end{equation}
The generalization of Eqs.~(\ref{CompToPau}) and (\ref{PauToComp}) read 
\begin{align}
\kket{\sigma_{\vec{k}\vec{l} }}  & = 
   \sum_{\vec{j}} (-1)^{\vec{j}\cdot\vec{k}} \ket{\vec{j}}\ket{\vec{j}+ \vec{l}(\text{mod}2)},\\
%
 \ket{\vec{l}}\ket{\vec{l}+ \vec{n}(\text{mod}2)} & = 
   \frac{1}{2^N}\sum_{\vec{m}} (-1)^{\vec{m}\cdot\vec{l}}\kket{\sigma_{\vec{m}\vec{n} }}.
\end{align}
% \afnote{Cambiar los $\mu$ y $\nu$ por indices latinos.}

### 2022-05-03 documentos/quantum_classical.tex
comentarios

\section{Quantum-Classical channels}
\label{quantum_classical}
\ddnote{Alguien revise este apendice, tiene solo una iterada}
A Quantum-classical (QC) channel is defined using an orthonormal basis in the Hilbert space. Let $B=\left\{ \ket{\psi_i} \right\}$ such a basis in $\hilbert$ with $\dim(\hilbert)=2^N$, the QC channel associated to $B$ is
\begin{equation}
\mcE^{\text{QC}}_B[\rho]=\sum_{i=1}^{2^N} \bra{\psi_i}\rho \ket{\psi_i} \projj{\psi_i},
\end{equation}
this is, QC channels project density matrices onto the corresponding diagonal matrix in the basis $B$~\cite{nathanson2007pauli}.

Consider now that the basis $B$ is the simultaneous eigenbasis of a maximal set of commuting Pauli strings denoted by $\text{set}(B)$, such set contains $2^N$ elements (including the identity), and there are $2^N+1$ of such sets~\cite{Lawrence2002}. We will prove that the QCs defined in this way are \pce{} channels with $2^N$ $1$s on their diagonal.

First we compute the components of $\mcE^{\text{QC}}_B$ in the basis $2^{-N/2}\left\{\sigma_{\vec \alpha}\right\}$,
\begin{equation}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=\frac{1}{2^N}\sum_{i=1}^{2^N} \bra{\psi_i}\sigma_{\vec k} \ket{\psi_i}\bra{\psi_i}\sigma_{\vec l} \ket{\psi_i}.
\end{equation}
To evaluate the components, observe that $\bra{\psi_i}\sigma_{\vec k} \ket{\psi_i}^2=1 \ \ \forall \sigma_{\vec k}\in \text{set}(B)$, and from the formula for the purity of $\ket{\psi_i}$,
\begin{equation}
1=\frac{1}{2^N}\sum_{\vec k} \bra{\psi_i} \sigma_{\vec k} \ket{\psi_i}^2,
\end{equation}
it follows that $\bra{\psi_i} \sigma_{\vec k'} \ket{\psi_i}=0 \ \ \forall \sigma_{\vec k'} \not\in \text{set}(B)$ since there only positive terms in the sum, and $2^N$ of them are already equal to $1$. Therefore,
\begin{align}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k}&=1,\nonumber\\ \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l'}&=\left(\mcE^{\text{QC}}_B\right)_{\vec l' \vec l'}=0 \ \ \forall \sigma_{\vec k} \in \text{set}(B) \ \ \forall \sigma_{\vec l'} \not\in \text{set}(B).
\end{align}
To compute $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}$ for $\sigma_{\vec k}, \sigma_{\vec l} \in \text{set}(B)$ with $\vec k \neq \vec l$, observe that $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=\left(\mcE^{\text{QC}}_B\right)_{\vec l \vec k}$, \ie{} the matrix corresponding to $\mcE^{\text{QC}}$ is an orthogonal projector. Thus, consider the block,
\begin{equation}
\left[ \begin{array}{cc}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k} & \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} \\ 
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} & \left(\mcE^{\text{QC}}_B\right)_{\vec l \vec l}
\end{array}  \right]=\left[ \begin{array}{cc}
1 & \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} \\ 
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} & 1
\end{array}  \right],
\end{equation}
it is easy to check that the latter is a projector only if $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=0$. Since there are exactly $2^N$ elements of the form $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k}$ with $\sigma_{\vec k} \in \text{set}(B)$, the channel $\mcE^{\text{QC}}_B$ is \pce{} with $2^N$ $1$s on its diagonal.

>>>>

\section{Quantum-Classical channels}
\label{quantum_classical}
A Quantum-classical (QC) channel is defined by using an orthonormal basis in the Hilbert space. Let $B=\left\{ \ket{\psi_i} \right\}$ such a basis in $\hilbert$ with $\dim(\hilbert)=2^N$, the QC channel associated to $B$ is
\begin{equation}
\mcE^{\text{QC}}_B[\rho]=\sum_{i=1}^{2^N} \bra{\psi_i}\rho \ket{\psi_i} \projj{\psi_i},
\end{equation}
this is, QC channels project density matrices onto the corresponding diagonal matrix in the basis $B$~\cite{nathanson2007pauli}.

Consider now that the basis $B$ is the simultaneous eigenbasis of a maximal set of commuting Pauli strings denoted by $\text{set}(B)$, such set contains $2^N$ elements (including the identity), and there are $2^N+1$ of such sets~\cite{Lawrence2002}. Now we proceed to demonstrate that the QCs defined in this way are \pce{} channels with $2^N$ $1$s on their diagonal.

First we compute the components of $\mcE^{\text{QC}}_B$ in the basis $2^{-N/2}\left\{\sigma_{\vec \alpha}\right\}$,
\begin{equation}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=\frac{1}{2^N}\sum_{i=1}^{2^N} \bra{\psi_i}\sigma_{\vec k} \ket{\psi_i}\bra{\psi_i}\sigma_{\vec l} \ket{\psi_i}.
\end{equation}
To evaluate the components, observe that $\bra{\psi_i}\sigma_{\vec k} \ket{\psi_i}^2=1 \ \ \forall \sigma_{\vec k}\in \text{set}(B)$, and from the formula for the purity of $\ket{\psi_i}$,
\begin{equation}
1=\frac{1}{2^N}\sum_{\vec k} \bra{\psi_i} \sigma_{\vec k} \ket{\psi_i}^2,
\end{equation}
it follows that $\bra{\psi_i} \sigma_{\vec k'} \ket{\psi_i}=0 \ \ \forall \sigma_{\vec k'} \not\in \text{set}(B)$ since there are only positive terms in the sum, and $2^N$ of them are already equal to $1$. Therefore,
\begin{align}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k}&=1,\nonumber\\ \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l'}&=\left(\mcE^{\text{QC}}_B\right)_{\vec l' \vec l'}=0 \ \ \forall \sigma_{\vec k} \in \text{set}(B) \ \ \forall \sigma_{\vec l'} \not\in \text{set}(B).
\end{align}
To compute $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}$ for $\sigma_{\vec k}, \sigma_{\vec l} \in \text{set}(B)$ with $\vec k \neq \vec l$, observe that $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=\left(\mcE^{\text{QC}}_B\right)_{\vec l \vec k}$, \ie{} the matrix corresponding to $\mcE^{\text{QC}}$ is an orthogonal projector. Thus, consider the block,
\begin{equation}
\left[ \begin{array}{cc}
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k} & \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} \\ 
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} & \left(\mcE^{\text{QC}}_B\right)_{\vec l \vec l}
\end{array}  \right]=\left[ \begin{array}{cc}
1 & \left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} \\ 
\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l} & 1
\end{array}  \right],
\end{equation}
it is easy to check that the latter is a projector only if $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec l}=0$. Since there are exactly $2^N$ elements of the form $\left(\mcE^{\text{QC}}_B\right)_{\vec k \vec k}$ with $\sigma_{\vec k} \in \text{set}(B)$, then the channel $\mcE^{\text{QC}}_B$ is \pce{} with $2^N$ $1$s on its diagonal.
